Software Clusterings with Vector Semantics and the Call Graph

### 2019-04-26 manuscript/manuscript.tex
Refactor manuscript and add ACDC MoJo Distance

In modern software systems, the need for adoption of modular software architectural schemas 
becomes inevident for robust maintainability of the overall software system.
Since the architecture of a software system is the most fundamental realization of a 
software system, when there is no description of the architecture of it, software 
architects usually make attempts to recover it. In the past, there have been attempts to 
identify the main components of a software system using hierarchical clustering 
algorithms \cite{maqbool_overview, limbo} mainly by clustering modules into subsystems, given 
its source code.    
The possibilities offered by machine learning techniques and practices in the Natural Language Processing (NLP) 
can also be applied on codebases as corpuses of text, thus enabling 
the extraction of useful information for the semantics of the source codebase 
inaugurating new modes of reviewing large amounts of source code.
An example is demonstrated through the Deep Code Search Project \cite{deepcodesearch} .

This paper aims to propose software clusterings using a combination of vector semantics on the source code provided by \emph{Doc2Vec} as well as the existing structure of 
the codebase via the call graph in order to form communities adn arrive at a module-level 
architecture viewed from the call graph perspective. 
We will study Linux codebase and we will compare it with ground truth in using the MoJo metric \cite{mojo} as well as 
compare it with other state-of-the-art and baseline clustering algorithms \cite{maqbool_overview, evaluation} in terms of 
\emph{stability}, \emph{authoritativeness} and \emph{extremity}

% XXX Not sure if needed

% We begin with a brief outline of the theoretical background concerning layered 
% architectures addressing its conjunction with software analysis tools and NLP practices. 
% Next, we propose methods for analyzing large codebases using techniques outlined in the 
% Methodology section. Next, we present the case study of analyzing the Linux Kernel 
% codebase using certain evaluation criteria, as well as results induced by the methodology 
% we have proposed. Finally, conclude and recommend roadmaps for future research on the 
% topic.

@@

\begin{table*}
  \caption{Experimental Results for Linux 4.21 Codebase}
    \label{tab:evaluation}
    \begin{tabular}{lrrrrrr}
    \toprule
    Algorithm & Number of Clusters & Cluster Size Range & Avg. Cluster & Standard Deviation & Median Cluster & MoJo Distance \\
    \midrule
    ACDC \cite{acdc} & 9055 & 1 -- 4245 & 4.53 & 45.82 & 2\\
    Average Linkage \cite{average} & 21 & 1--3406 & 163 & 725.11 & 1 & 2092 \\
    Complete Linkage \cite{complete} & 21 & 1--1529 & 163 & 412.34 & 19.0 & 1710 \\
    LIMBO \cite{limbo} & & \\
    SADE & 10 (\pm 2)  & 2 (\pm 0) -132 (\pm 13) & 64 (\pm 4) & 40 (\pm 4) & 65 (\pm 10) & 243 (\pm 1)  \\
    SADE (Directed) & 5 (\pm 2) & 1 (\pm 1) - 614 (\pm 1) & 141 (\pm 39) & 253 (\pm 25) & 2 (\pm 0.3)  & 237 (\pm 2) \\

    Ward Linkage \cite{ward} & 21 & 21--948 & 163 & 223 & 70.0 & 1138 \\
    Ground Truth & 21 & 1--1348 & 163 & 340.62 & 11.0 & -- \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

For the preprocessing stage use the NLP package spaCy \cite{spacy} which provided us with very useful tools for the extraction. After preprocessing, we use gensim \cite{gensim} to train the Doc2Vec model. First of all, we tokenize the code and remove the stopwords. After that, for each token, we split it into its constituent parts using dynamic programming and lemmatize 
each individual sub-token. 
For example the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, \texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.

% The identifier \texttt{inprogress} should be split into \texttt{in} and \texttt{progress} 
% which provides us with a more meaningful content for the word embeddings from a 
% linguistic point of view. The identifier segmentation problem is reduced to using an arbitrary number of separators we must split the identifier $s = i_1 i_2 \dots i_n$ such
% that the total cost of splitting such words minimizing the split cost. Using a word-level $n$-gram model we define the cost function of a (sub)-word $s_k$ given prior words $s_{k-1}, s_{k-2} \dots, s_{k-n + 1}$ and a dataset $\mathcal D$ as $c(s_k | s_{k-1} \dots s_{k - n + 1}, \mathcal D) = - \log \hat p_{MLE} (s_k | s_{k-1} \dots s_{k-n+1}, \mathcal D)$ where $\hat p_{MLE} (\cdot)$ is the Maximum Likelihood Estimator of the event. Let $L(m)$ be the cost of the optimal split in the $m$ character of the identifier. Then

% $$L(m) = 
% \begin{cases}
% 	\min_{0 \le \ell < m} \{ L(\ell) + c(s[\ell:m]  \mid \\ \text{ prior words to } s% [\ell:m] ) \} & 1 \le m \le |s| \\
%	0 & m = 0

%\end{cases}
%$$

@@

\paragraph{Combining semantics and the Call Graph} The input to the clustering procedure
 is the call graph with cosine similarities of the various modules as weights, thus forming a
  weighted directed graph. Then we run the Louvain Clustering Algorithm \cite{louvain} in
   order to obtain the software clustering via maximizing the modularity function with a
    greedy approach.

    %Further refining can also be done by re-running the algorithm on the contracted graph with mean embeddings as embeddings for each node.

@@

% used to handle graphs and graph operations was NetworkX \cite{nx}. The code is available on GitHub.

\paragraph{Decision Rationale for Evaluation System} We have chosen to use Linux as a codebase to evaluate our method on since it is a large and complex system with 27 years of continuous development, spanning 20.3 millions SLOC at the time of writing.  Moreover, it is easy to find ground truth system due to its clear directory structure and construct testcases for
evaluations. Finally, it was also used in related studies \cite{acdc, evaluation} as a reference system for evaluation.

\paragraph{Experimental Setting} Our method was run on Linux 4.21,
 consisting of 20.3 million SLOC against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} and Ward-Linkage \cite{ward} as well as ACDC and LIMBO. The results are presented in Table \ref{tab:evaluation}. As ground truth we have used the first level directories as a target clustering and as input we have considered the modules of the one-top directories. For example, the source code file \texttt{drivers/net/ieee802154/mcr20a.c} has a ground truth value of \texttt{drivers} and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. As a clustering distance metric we have used the MoJo metric, as proposed in reference \cite{mojo}. Since, Louvain Community Detection produces 
 different (but very similar) results every time it runs, we ran the simulations multiple times and averaged the results.

\section{Results, Evaluation, Discussion}

@@

% TODO Needs review

% The architecture of a software system is a \emph{"living organism"} meaning that it is 
% changing constantly over time, even for small projects. The evolution software 
% architecture may lead to architectural decay \cite{large_study}. The identification of such 
% a decay is non-trivial and requires a large amount of effort and understanding of the 
% software system in order to be diagnosed.

% Architectural recovery is an active research 
%field aiming to make software systems more understandable and maintainable over time. The 
%study of architectural change requires that the architecture at a given point in time, 
%for example after a major release, during a system's evolution must be inferred. 
%Various algorithms have been proposed for realizing architectural schemata many of which use 
%different features for describing the source code.

>>>>

In modern software systems, the need for adoption of modular software architectural schemata 
becomes inevitable for robust maintainability of the overall software system.
Since the architecture of a software system is the most fundamental realization of a 
software system, when there is no description of the architecture of it, software 
architects usually make attempts to recover it. In the past, there have been attempts to 
identify the main components of a software system using hierarchical clustering 
algorithms \cite{maqbool_overview, limbo} mainly by clustering modules into subsystems, given 
its source code.    
The possibilities offered by machine learning techniques and practices in the Natural Language Processing (NLP) 
can also be applied on codebases as corpora of text, thus enabling 
the extraction of useful information for the semantics of the source codebase 
inaugurating new modes of reviewing large amounts of source code.
An example is demonstrated through the Deep Code Search Project \cite{deepcodesearch} .

This paper aims to propose software clusterings using a combination of vector semantics on the source code provided 
by \emph{Doc2Vec} as well as the existing structure of 
the codebase via the call graph in order to form communities and arrive at a module-level 
architecture viewed from the call graph perspective. 
We will study Linux codebase and we will compare it with ground truth in using the MoJo metric \cite{mojo} as well as 
compare it with other state-of-the-art and baseline clustering algorithms \cite{maqbool_overview, evaluation} in terms of 
\emph{stability}, \emph{authoritativeness} and \emph{extremity}

@@

\begin{table*}
  \caption{Experimental Results for Linux 4.21 Codebase}
    \label{tab:evaluation}
    \begin{tabular}{lrrrrrr}
    \toprule
    Algorithm & Number of Clusters & Cluster Size Range & Avg. Cluster & Standard Deviation & Median Cluster & MoJo Distance \\
    \midrule
    ACDC \cite{acdc} & 9055 & 1 -- 4245 & 4.53 & 45.82 & 2 & 33694\\
    Average Linkage \cite{average} & 21 & 1--3406 & 163 & 725.11 & 1 & 2092 \\
    Complete Linkage \cite{complete} & 21 & 1--1529 & 163 & 412.34 & 19.0 & 1710 \\
    Ward Linkage \cite{ward} & 21 & 21--948 & 163 & 223 & 70.0 & 1138 \\

    LIMBO \cite{limbo} & 21 & & \\
    SADE & 10 (\pm 2)  & 2 (\pm 0) -132 (\pm 13) & 64 (\pm 4) & 40 (\pm 4) & 65 (\pm 10) & 243 (\pm 1)  \\
    SADE (Directed) & 5 (\pm 2) & 1 (\pm 1) - 614 (\pm 1) & 141 (\pm 39) & 253 (\pm 25) & 2 (\pm 0.3)  & 237 (\pm 2) \\

    Ground Truth & 21 & 1--1348 & 163 & 340.62 & 11.0 & -- \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

For the prepossessing stage use the NLP package spaCy \cite{spacy} which provided us with very useful tools for the extraction. 
After prepossessing, we use Gensim \cite{gensim} to train the Doc2Vec model. 
First of all, we tokenize the code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming and lemmatize 
each individual sub-token. 
For example the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.

@@

\paragraph{Combining semantics and the Call Graph} 
The input to the clustering procedure
is the call graph with cosine similarities of the various modules as weights, thus forming a
weighted directed graph. Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
In case we want to use edge directionality we do a bipartite transformation, detect communities and
merge the results with a disjoint-set data structure.

@@

\paragraph{Decision Rationale for Evaluation System} 
We have chosen to use Linux as a codebase to evaluate our method on 
since it is a large and complex system with 27 years of continuous development, spanning 20.3 millions source lines of code (SLOC) at the time of writing.  
Moreover, it is easy to find ground truth system due to its clear directory structure and construct test-cases for
evaluations. 
Finally, it was also used in related studies \cite{acdc, evaluation} as a reference system for evaluation.

\paragraph{Experimental Setting} Our method was run on Linux 4.21,
consisting of 20.3 million SLOC against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} 
and Ward-Linkage \cite{ward} as well as ACDC and LIMBO ($B = 10, S = \infty$). The results are presented in Table \ref{tab:evaluation}. 
As ground truth we have used the first level directories as a target clustering and as input we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. 
As a clustering distance metric we have used the MoJo metric, as proposed in reference \cite{mojo}. 
Since, Louvain Community Detection produces  different (but very similar) results every time it runs, we ran the simulations 
multiple times and averaged the results.

\section{Results, Evaluation \& Discussion}

@@
